---
title: "Causal inference"
author: "Daoming Wang"
date: "2021/7/31"
output:
  html_document: 
    theme: flatly
    highlight: espresso
    toc: true
    toc_depth: 4
    toc_float: true
  word_document: default
  pdf_document:
    includes:
      in_header: header.tex
      keep_tex: yes
      latex_engine: xelatex
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Preparation

### 1.1 Import

Import packages and functions.

```{r 1.1, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
source("functions.R")
```

### 1.2 Inputs

Read input files.

```{r 1.2, echo=TRUE}
load("01.cleanData/SV_lld/dsgv_lld.RData")
load("01.cleanData/SV_lld/vsgv_lld.RData")

info    <- read.table("01.cleanData/SV_info/20200801_LLD_300OB_Informative_species_information.tsv",
                      sep = "\t", header = T, stringsAsFactors = F)
vsv_info<-read.table("01.cleanData/SV_info/20200801_LLD_300OB_vsgv_info_anno.tsv",
                     sep = "\t",header = T,stringsAsFactors = F, quote = "")
dsv_info<-read.table("01.cleanData/SV_info/20200801_LLD_300OB_dsgv_info_anno.tsv",
                     sep = "\t",header = T,stringsAsFactors = F,quote = "")


lld_ba <- read.table("01.cleanData/phen_lld/20200801_LLD_39BA_1135samples.tsv")
lld_basic <- read.table("01.cleanData/phen_lld/20200801_LLD_basic_1135samples.tsv")
lld_phen<-read.table("00.rawData/LLD_phenotype/20200226_207allPhenotypes_1135patients.txt")

lld_exp<-lld_phen[,c(81:207)]

lld_abun<-read.table("01.cleanData/mbio_lld/LLD_SV_55species_abun_1135samples.tsv", check.names = F) 

vsv_ba_lm_adjAbunPCs_res.sig.anno.edge<-read.table("09.Microbial_GWAS/vsv_ba_lm_adjAbunPCs.sig.anno.final.tsv", 
                                                sep = "\t", header = T, check.names = F)
dsv_ba_lm_adjAbunPCs_res.sig.anno.edge<-read.table("09.Microbial_GWAS/dsv_ba_lm_adjAbunPCs.sig.anno.final.tsv", 
                                                sep = "\t", header = T, check.names = F)

load("01.cleanData/SV_lld/lld_msv_pc_cum0.6.RData")
```

### 1.3 Preparation

```{r 1.3}
## Prepare covariate table
# lld
lld_abun$Others<-1-rowSums(lld_abun)
lld_abun_clr<-abundances(x=as.data.frame(na.omit(lld_abun)), transform="clr") %>%as.data.frame
lld_abun_clr <- lld_abun_clr[match(rownames(lld_abun), rownames(lld_abun_clr)),]
rownames(lld_abun_clr) <- rownames(lld_abun)

lld_covar<-cbind(lld_basic,lld_abun_clr,lld_msv_pc_cum0.6)

covar <- c('Gender','Age','BMI','Reads_number')

if (!dir.exists("11.Causal_inference")) {dir.create("11.Causal_inference")}
if (!dir.exists("11.Causal_inference/RData")) {dir.create("11.Causal_inference/RData")}

lld_assoc_vsv <- vsv_ba_lm_adjAbunPCs_res.sig.anno.edge$SV %>% unique
lld_assoc_dsv <- dsv_ba_lm_adjAbunPCs_res.sig.anno.edge$SV %>% unique

vsgv_lld<-vsgv_lld[,lld_assoc_vsv]
dsgv_lld<-dsgv_lld[,lld_assoc_dsv]

```

## 2 Associations

### 2.1 Associations between SV and exposure

#### 2.1.1 Associations between vSV and exposure

```{r 2.1.1, eval=FALSE}
lld_vsv_exp_lm_adjAbunPCs_res<-lm_btw_mats_adjAbunPCs2(vsgv_lld,lld_exp,lld_basic,covar,lld_abun_clr, lld_msv_pc_cum0.6,info, 9)
save(lld_vsv_exp_lm_adjAbunPCs_res, file = "11.Causal_inference/RData/lld_vsv_exp_lm_adjAbunPCs_res.RData")
```

#### 2 Associations between dSV and exposure

```{r 2.1.2, eval=FALSE}
## Prepare covariate table
lld_dsv_exp_lr_adjAbunPCs_res<-lr_btw_mats_adjAbunPCs2(dsgv_lld,lld_exp,lld_basic,covar,lld_abun_clr, lld_msv_pc_cum0.6,info,9)
save(lld_dsv_exp_lr_adjAbunPCs_res, file = "11.Causal_inference/RData/lld_dsv_exp_lr_adjAbunPCs_res.RData")
```


### 2.2 Associations between BA and exposure

```{r 2.2, eval=FALSE}
lld_ba_exp_lm_res <- lm_btw_mats(lld_ba, lld_exp, lld_basic, covar)
save(lld_ba_exp_lm_res, file = "11.Causal_inference/RData/lld_ba_exp_lm_res.RData")
```


## 3 Get variable groups with potential mediation effect

### 3.1 vSV

```{r 3.1, eval=FALSE}
## load SV-BA associations restuls
vsv_ba_lm_res.sig.anno.edge<-read.table("09.Microbial_GWAS/vsv_ba_lm_adjAbunPCs.sig.anno.tsv",
                                        sep = "\t", header = T,stringsAsFactors = F,check.names = F, quote = "")

load("11.Causal_inference/RData/lld_vsv_exp_lm_adjAbunPCs_res.RData")
load("11.Causal_inference/RData/lld_ba_exp_lm_res.RData")

lld_vsv_exp_lm_adjAbunPCs.edge <- lld_vsv_exp_lm_adjAbunPCs_res
lld_vsv_exp_lm_adjAbunPCs.edge$Y<-as.character(lld_vsv_exp_lm_adjAbunPCs.edge$Y)
lld_vsv_exp_lm_adjAbunPCs.edge$X<-as.character(lld_vsv_exp_lm_adjAbunPCs.edge$X)

lld_ba_exp_lm_adjAbunPCs.edge  <- lld_ba_exp_lm_res
lld_ba_exp_lm_adjAbunPCs.edge$Y<-as.character(lld_ba_exp_lm_adjAbunPCs.edge$Y)
lld_ba_exp_lm_adjAbunPCs.edge$X<-as.character(lld_ba_exp_lm_adjAbunPCs.edge$X)

## Get candidate groups
exp_id <- colnames(lld_exp)
vsv_id <-vsv_ba_lm_res.sig.anno.edge$SV %>%
  as.character(.) %>%
  .[!duplicated(.)] %>%
  sort(.,decreasing = F)
ba_id  <- vsv_ba_lm_res.sig.anno.edge$BA %>%
  as.character(.) %>%
  .[!duplicated(.)]

exp_vsv_ba_df <- NULL

for (i in exp_id) {
  cat(paste(i,"\n"))
  for (j in vsv_id) {
    for (k in ba_id) {
      #i<-exp_id[1]
      #j<-vsv_id[1]
      #k<-ba_id[1]
      
      p_vsv_exp <- lld_vsv_exp_lm_adjAbunPCs.edge$p[lld_vsv_exp_lm_adjAbunPCs.edge$X==i & lld_vsv_exp_lm_adjAbunPCs.edge$Y==j]
      p_ba_exp  <- lld_ba_exp_lm_adjAbunPCs.edge$p[  lld_ba_exp_lm_adjAbunPCs.edge$X==i & lld_ba_exp_lm_adjAbunPCs.edge$Y==k]
      p_ba_vsv  <- vsv_ba_lm_res.sig.anno.edge$LLD.p[vsv_ba_lm_res.sig.anno.edge$SV==j & vsv_ba_lm_res.sig.anno.edge$BA==k]
      
      if(length(p_ba_vsv) > 0& !is.na(p_ba_vsv) & !is.na(p_vsv_exp)  & !is.na(p_ba_exp) ){
        if(p_vsv_exp<0.05 & p_ba_exp<0.05 & p_ba_vsv<0.05){
        exp_vsv_ba_df <- rbind(exp_vsv_ba_df,c(i, j, k))
        }
      }
    }
  }
}

save(exp_vsv_ba_df,file = "11.Causal_inference/RData/exp_vsv_ba_df_model2.RData")
if (!dir.exists("11.Causal_inference/")) {dir.create("11.Causal_inference/")}
write.table(exp_vsv_ba_df, "11.Causal_inference/exp_vsv_ba_df_model2.tsv",
            sep = "\t", col.names = T, row.names = F, quote = F)
```


### 3.2 dSV

```{r 3.2, eval=FALSE}
## load SV-BA associations restuls
dsv_ba_lm_res.sig.anno.edge<-read.table("09.Microbial_GWAS/dsv_ba_lm_adjAbunPCs.sig.anno.tsv",
                                        sep = "\t", header = T,stringsAsFactors = F,check.names = F, quote = "")

load("11.Causal_inference/RData/lld_dsv_exp_lr_adjAbunPCs_res.RData")
load("11.Causal_inference/RData/lld_ba_exp_lm_res.RData")

lld_dsv_exp_lr_adjAbunPCs.edge <- lld_dsv_exp_lr_adjAbunPCs_res
lld_dsv_exp_lr_adjAbunPCs.edge$Y<-as.character(lld_dsv_exp_lr_adjAbunPCs.edge$Y)
lld_dsv_exp_lr_adjAbunPCs.edge$X<-as.character(lld_dsv_exp_lr_adjAbunPCs.edge$X)

lld_ba_exp_lm_adjAbunPCs.edge  <- lld_ba_exp_lm_res
lld_ba_exp_lm_adjAbunPCs.edge$Y<-as.character(lld_ba_exp_lm_adjAbunPCs.edge$Y)
lld_ba_exp_lm_adjAbunPCs.edge$X<-as.character(lld_ba_exp_lm_adjAbunPCs.edge$X)

## Get candidate groups
exp_id <- colnames(lld_exp)
dsv_id <-dsv_ba_lm_res.sig.anno.edge$SV %>%
  as.character(.) %>%
  .[!duplicated(.)] %>%
  sort(.,decreasing = F)
ba_id  <- dsv_ba_lm_res.sig.anno.edge$BA %>%
  as.character(.) %>%
  .[!duplicated(.)]

exp_dsv_ba_df <- NULL

for (i in exp_id) {
  cat(paste(i,"\n"))
  for (j in dsv_id) {
    for (k in ba_id) {
#      i<-exp_id[1]
#     j<-dsv_id[1]
#      k<-ba_id[1]
      
      p_dsv_exp <- lld_dsv_exp_lr_adjAbunPCs.edge$p[lld_dsv_exp_lr_adjAbunPCs.edge$Y==i & lld_dsv_exp_lr_adjAbunPCs.edge$X==j]
      if(length(p_dsv_exp)==0){p_dsv_exp<-1}
      p_ba_exp  <- lld_ba_exp_lm_adjAbunPCs.edge$p[  lld_ba_exp_lm_adjAbunPCs.edge$X==i & lld_ba_exp_lm_adjAbunPCs.edge$Y ==k]
      p_ba_dsv  <- dsv_ba_lm_res.sig.anno.edge$LLD.p[dsv_ba_lm_res.sig.anno.edge$SV==j & dsv_ba_lm_res.sig.anno.edge$BA==k]
      if(length(p_ba_dsv)==0){p_ba_dsv<-1}
      
      if(length(p_ba_dsv) > 0 & !is.na(p_ba_dsv) & length(p_dsv_exp) > 0 & !is.na(p_dsv_exp) & length(p_ba_exp) > 0  & !is.na(p_ba_exp) ){
        if(p_dsv_exp<0.05 & p_ba_exp<0.05 & p_ba_dsv<0.05){
        exp_dsv_ba_df <- rbind(exp_dsv_ba_df,c(i, j, k))
        }
      }
    }
  }
}

save(exp_dsv_ba_df,file = "11.Causal_inference/RData/exp_dsv_ba_df.RData")
if (!dir.exists("11.Causal_inference/")) {dir.create("11.Causal_inference/")}
write.table(exp_dsv_ba_df, "11.Causal_inference/exp_dsv_ba_df.tsv",
            sep = "\t", col.names = T, row.names = F, quote = F)
```


## 4 Mediation analysis

### 4.1 vSV

```{r 4.1, eval=FALSE}
load("11.Causal_inference/RData/exp_vsv_ba_df_model2.RData")

covar_abun<-str_replace_all(exp_vsv_ba_df[,2],"\\:\\d+_\\d+.*","") 
exp_vsv_ba_df<-cbind(exp_vsv_ba_df, covar_abun)

date() # "Wed Nov 11 21:46:05 2020"
vsv_mediation_res<-NULL
for (i in 1:nrow(exp_vsv_ba_df)) {
  #i<-1131
  cat(paste(i,"\n"))
  res_i<-lm_bimediation(exp_vsv_ba_df[i,],
                        indvDf=lld_exp, 
                        dvDf1=vsgv_lld, 
                        dvDf2=lld_ba, 
                        covDf=lld_covar, 
                        covar=c('Gender','Age','BMI','Reads_number'))
  
  vsv_mediation_res[[i]]<-res_i
}
date()


vsv_mediation <- matrix(unlist(vsv_mediation_res), ncol = 34, byrow = T) %>% as.data.frame
colnames(vsv_mediation)<-names(vsv_mediation_res[[1]])

save(vsv_mediation, file = "11.Causal_inference/RData/vsv_mediation_model2.RData")
write.table(vsv_mediation,"11.Causal_inference/vsv_mediation_model2.tsv",
            sep = "\t", col.names = T, row.names = F, quote = F)

```


### 4.2 dSV

```{r 4.2, eval=FALSE}
load("11.Causal_inference/RData/exp_dsv_ba_df.RData")

covar_abun<-str_replace_all(exp_dsv_ba_df[,2],"\\:\\d+_\\d+.*","") %>% match(.,info$organism) %>% info$Metaphlan3_name[.]
covar_abun[info$Abundance_available[match(covar_abun, info$Metaphlan3_name)]=='No']<-NA
exp_dsv_ba_df<-cbind(exp_dsv_ba_df, covar_abun)

date() # "Fri Nov 13 15:29:51 2020"
dsv_mediation_res<-NULL
for (i in 1:nrow(exp_dsv_ba_df)) {
  #i<-100
  cat(paste(i,"\n"))
  

  res_i<-lm_bimediation(exp_dsv_ba_df[i,],
                        indvDf=lld_exp, 
                        dvDf1=dsgv_lld, 
                        dvDf2=lld_ba, 
                        covDf=lld_covar, 
                        covar=c('Gender','Age','BMI','Reads_number'))
  
  dsv_mediation_res[[i]]<-res_i
}
date()

dsv_mediation <- matrix(unlist(dsv_mediation_res), ncol = 34, byrow = T) %>% as.data.frame
colnames(dsv_mediation)<-names(dsv_mediation_res[[1]])

save(dsv_mediation, file = "11.Causal_inference/RData/dsv_mediation.RData")
write.table(dsv_mediation,"11.Causal_inference/dsv_mediation.tsv",
            sep = "\t", col.names = T, row.names = F, quote = F)

```


## 5 Visualization
### 5.1 Prepare final table

```{r 5.1}
load("11.Causal_inference/RData/exp_vsv_ba_df_model2.RData")
#load("11.Causal_inference/RData/dsv_mediation.RData")

vsv_mediation[,c(5:34)]<-as.data.frame(sapply(vsv_mediation[,c(5:34)], as.numeric))
#dsv_mediation[,c(5:34)]<-as.data.frame(sapply(dsv_mediation[,c(5:34)], as.numeric))



vsv_mediation_final<-vsv_mediation
vsv_mediation_final$dir1.ACME.fdr<-p.adjust(vsv_mediation$dir1.ACME.p, method = 'fdr')
vsv_mediation_final$dir2.ACME.fdr<-p.adjust(vsv_mediation$dir2.ACME.p, method = 'fdr')

#dsv_mediation_final<-dsv_mediation
#dsv_mediation_final$dir1.ACME.fdr<-p.adjust(dsv_mediation$dir1.ACME.p, method = 'fdr')
#dsv_mediation_final$dir2.ACME.fdr<-p.adjust(dsv_mediation$dir2.ACME.p, method = 'fdr')

#mediation_final<-rbind(vsv_mediation_final, dsv_mediation_final)
mediation_final<-vsv_mediation_final
sv_type<-c(rep('vSV', nrow(vsv_mediation_final)))
mediation_final<-data.frame(sv_type, mediation_final)

mediation_final[,c(5:34)]<-as.data.frame(sapply(mediation_final[,c(5:34)], as.numeric))
mediation_final$MediationDirection<-rep('none', length(mediation_final$MediationDirection))

fdr.cutoff<-0.05
#mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr<fdr.cutoff &
#                                         mediation_final$dir2.ACME.fdr<fdr.cutoff] <- 'both'
#mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr<fdr.cutoff &
#                                         mediation_final$dir2.ACME.fdr>=fdr.cutoff] <- 'indv_dv1_dv2'
#mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr>fdr.cutoff &
#                                         mediation_final$dir2.ACME.fdr<=fdr.cutoff] <- 'indv_dv2_dv1'

mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr<fdr.cutoff & 
                                     mediation_final$dir1.Prop.mediated.p<0.05 &
                                     mediation_final$dir2.ACME.fdr<fdr.cutoff &
                                     mediation_final$dir2.Prop.mediated.p<0.05] <- 'both'
mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr<fdr.cutoff &
                                     mediation_final$dir1.Prop.mediated.p<0.05 &
                                     mediation_final$dir2.ACME.fdr>fdr.cutoff] <- 'indv_dv1_dv2'
mediation_final$MediationDirection[mediation_final$dir1.ACME.fdr>fdr.cutoff &
                                     mediation_final$dir2.ACME.fdr<fdr.cutoff &
                                     mediation_final$dir2.Prop.mediated.p<0.05] <- 'indv_dv2_dv1'

table(mediation_final$MediationDirection,mediation_final$sv_type)



save(mediation_final, file = "11.Causal_inference/RData/mediation_final_model2.RData")
write.table(mediation_final, "11.Causal_inference/mediation_model2.tsv",
            sep = "\t", col.names = T, row.names = F, quote = F)
```

### 5.2 Venn diagram

```{r 5.2}

## venn diagram
mediation_final_count<-table(mediation_final$MediationDirection)
pdf("11.Causal_inference/mediation_count.venn.pdf", width = 2, height = 2)
draw.pairwise.venn(mediation_final_count[2]+mediation_final_count[1],
                   mediation_final_count[3]+mediation_final_count[1],
                   mediation_final_count[1], 
                   category = c("Direction 1", "Direction 2"), lty = rep("blank",2), fill =c("#4472c4", "#00b050"), alpha = rep(0.5, 2), cat.pos = c(0, 0), cat.cex = c(0.5, 0.5),cat.dist = rep(0.025, 2), scaled = F)
dev.off()
```



### 5.3 Sankey diagram of direction 1

```{r 5.3}
mediation_final<-read.table("11.Causal_inference/mediation.tsv", 
                            sep = "\t", header = T, stringsAsFactors = F)
mediation_plot_dir1<-mediation_final[mediation_final$MediationDirection=='indv_dv1_dv2' &
                                       mediation_final$dir1.Prop.mediated.p < 0.05 &
                                       mediation_final$dir1.Prop.mediated > 0.25,]#mediation_final$dir1.Prop.mediated>0.2

mediation_plot_dir1$fre<-1
sankey_colors<-c(wes_palette("Darjeeling1", length(unique(mediation_plot_dir1$dv2)), type = "continuous"))

#mediation_plot_dir1$indv<-factor(mediation_plot_dir1$indv, levels = c("fruits", "pasta" , "sugar_sweets","how_often_breakfast", "how_often_fruits","how_often_fish"  ,"how_often_alcohol" ,           "red_wine",                     "white_wine"        ,"how_often_muesli"  ,           "how_often_boiled_potatos"   ,  "how_often_soda","how_often_chocolade",   "how_often_nuts" ,           "smk_current","beta_sympathomimetic_inhaler" ,"oral_contraceptive"   ,        "how_often_pizza" ))

pdf("11.Causal_inference/mediation_dir1.sankey.pdf",width = 7,height = 3.5)
ggplot(mediation_plot_dir1,aes(axis1 = mediation_plot_dir1$indv, 
                               axis2 = mediation_plot_dir1$dv1, 
                               axis3 = mediation_plot_dir1$dv2,
                               y= mediation_plot_dir1$fre))+
  scale_x_discrete(limits = c("Lifestyle factors", "SVs", "BAs")) +
  geom_alluvium(aes(fill = mediation_plot_dir1$dv2),alpha = .55)+
  geom_stratum(alpha = 0,color = adjustcolor( "white", alpha.f = 1),size=1.2)+
  geom_text(stat = "stratum",cex=2,aes(label = after_stat(stratum))) +
  scale_color_manual(values=sankey_colors)+
  scale_fill_manual(values=sankey_colors)+
  theme_minimal()+
  theme(legend.position="none",
        axis.text = element_text(size = 10),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid=element_blank())
dev.off()

```


### 5.4 Sankey diagram of direction 2

```{r 5.4}
mediation_final<-read.table("11.Causal_inference/mediation.tsv", 
                            sep = "\t", header = T, stringsAsFactors = F)
mediation_plot_dir2<-mediation_final[mediation_final$MediationDirection=='indv_dv2_dv1' &
                                       mediation_final$dir2.Prop.mediated.p < 0.05,]#mediation_final$dir2.Prop.mediated>0.2

mediation_plot_dir2$fre<-1
sankey_colors<-c(wes_palette("Darjeeling1", length(unique(mediation_plot_dir2$dv1)), type = "continuous"))

pdf("11.Causal_inference/mediation_dir2.sankey.pdf",width = 7,height = 4)
ggplot(mediation_plot_dir2,aes(axis1 = mediation_plot_dir2$indv, 
                               axis2 = mediation_plot_dir2$dv2, 
                               axis3 = mediation_plot_dir2$dv1,
                               y= mediation_plot_dir2$fre))+
  scale_x_discrete(limits = c("Lifestyle factors", "BAs", "SVs")) +
  geom_alluvium(aes(fill = mediation_plot_dir2$dv1),alpha = .55)+
  geom_stratum(alpha = 0,color = adjustcolor( "white", alpha.f = 0.5),size=0.5)+
  geom_text(stat = "stratum",cex=2,aes(label = after_stat(stratum))) +
  scale_color_manual(values=sankey_colors)+
  scale_fill_manual(values=sankey_colors)+
  theme_minimal()+
  theme(legend.position="none",
        axis.text = element_text(size = 10),
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        panel.grid=element_blank())
  
dev.off()

```
