---
title: "Metabolomics"
author: "Arnau Vich"
date: "01/2020"
output: html_document
---

Introduction
===

In this project we aim to investigate the relation between fecal metabolites and inflammatory bowel diseases phenotypes. The cohort consist of 750 samples of different patients, devided in 255 controls from the population cohort LifeLines (NL), 270 patients with Crohns disease and 211 patients with ulcerative colitis from the 1000IBD cohort (NL). 

Metabolites were measured using METABOLON platform combining untargeted metabolite detection with short chain fatty acids measurments. Untargeted metabolite detection consisted on an ultrahigh performance liquid chromatography tandem mass spectrocopy (UPLC-MS/MS): 2 acidic positive ion condition (C18 columns) + 1 Basic negative ion optimazed (C18 column) + 1 Basic negative ionization (Hilic column).

Metabolites will be combined with the following data layers: matching shot-gun sequencing metagenomics (from the same sample), host phenotypes at time of sampling (including disease phenotypes), diet (food frequency questionnaires) & host genetics (GSA and WES data). 

Load functions
===

```{r, warning=F,message=FALSE, echo=TRUE}
library(corrplot)
library(plotly)
library(pheatmap)
library(NbClust)
library(reshape2)
library(ggrepel)
library(dplyr)
library(eulerr)
library(vegan)
library(ape)
library(ggridges)
library (UpSetR)
library(ggplot2)
library(psych)
library(ggord)
library(mice)
library(caret)
library(factoextra)
library(Rtsne)
library(heatmaply)
library(ggpubr)
library(ggbiplot)
library(coin)
library(DT)
library(pvclust)
library(compositions)
library(RColorBrewer)

#Inverse rank transformation function. 
invrank= function(x) {qnorm((rank(x,na.last="keep")-0.5)/sum(!is.na(x)))}

`%ni%` <- Negate(`%in%`)
```

Import data
---

a) Metabolites: Here we use both raw values provided by Metabolon (metabolites AUC) and transformed values. Transformation consist of two main procedures: adjusting medians of each metabolite to 1 (making distributions more symetric) and imputing missing values (missing values are replaced by the minimum recorded value of each metabolite). Raw values will be used to calculate prevalence and filter metabolites with high percentange of NA's

b) Short Chain Fatty Acids (SFCA): Together with the untargetted metabolite assessment, SCFA concentrations were measured in the same fecal samples (μg/g)

c) Host phenotypes: Information about the host (age,sex,BMI,etc.). 

d) Experiment variables: Different technical variables from the untargetted metabolic experiments. We will use this information as a potential counfounding factors. 

e) Faecal characteristics: Two phenotypes that describe the intestinal transit time and the sample moisture: bowel movement a day and stool consistency (soft or hard). Bristol stool scale was not available for the patients with IBD, however, this charactersitics were estimated from the harvey bradshaw score (HBI) and other clinical scores.

f) Food: Diet information captured from food frequency questionnaires (FFQs). Each element has been adjusted for nutrient density.

g) Metagenomics taxa: MetaPhlan's profiles (v2.7)

```{r, warning, message=F, echo=T}
#transformed metabolites (raw data transformed to adjust distrubution, medians = 1)
all_metabolites <- read.delim("~/Documents/Project_Metabolome/1.Input_files/all_metabolites2.txt", row.names=1)

#raw values (AUC of peaks)
all_metabolites_raw <- read.delim("~/Documents/Project_Metabolome/1.Input_files/all_metabolites_raw2.txt", row.names=1)

#Metabolites annotation
annot <- read.delim("~/Documents/Project_Metabolome/1.Input_files/info_metabolites_v2.txt", row.names=1)
annot$SUPER.PATHWAY.1=NULL

#Phenotypes

phenos_ibd=read.delim("~/Documents/Project_Metabolome/1.Input_files/Merged_IBD_phenos_clean_v2.txt",row.names = 1)
phenos_lld=read.delim("~/Documents/Project_Metabolome/1.Input_files/Merged_LLD_phenos_clean_v2.txt", row.names=1)
cc_pheno=read.delim("~/Documents/Project_Metabolome/1.Input_files/mini_pheno.txt", row.names=1)

# Storage

#translate id's 
IDs_LLD <- read.delim("~/Documents/Project_Metabolome/1.Input_files/IDs_LLD.txt", header=FALSE)
IDs_IBD <- read.delim("~/Documents/Project_Metabolome/1.Input_files/IDs_IBD.txt", header=FALSE)

#Metagenomics taxa

IBD_taxa <- read.delim("~/Documents/Project_Metabolome/2.taxa/IBD_taxonomy_unstrat_clean.txt", row.names=1)
LLD_taxa <- read.delim("~/Documents/Project_Metabolome/2.taxa/LLD_taxonomy_unstrat_clean.txt", row.names=1)


```


Describe the categories of metabolites
---

```{r, warning=F,message=FALSE, echo=F}
met_2=as.data.frame(table(annot$SUPER.PATHWAY))
met_1=as.data.frame(table(annot$SUB.PATHWAY))

yy=ggplot(met_2, aes(x = reorder(Var1, -Freq),Freq)) + geom_bar(stat = "identity", color="black") + theme_classic() + theme(axis.title.x = element_blank(),axis.text.y = element_text(size = 12)) + coord_flip() + xlab("") + ylab("Number of metabolites")

ggplotly(yy)
```

```{r, warning=F,message=FALSE, echo=F,fig.height=15,}
xx=ggplot(met_1, aes(x = reorder(Var1, -Freq),Freq)) + geom_bar(stat = "identity", color="black") + theme_classic() + theme(axis.title.x = element_blank(),axis.text.y = element_text(size = 7)) + coord_flip() + xlab("") + ylab("Number of metabolites")

ggplotly(xx)
```


Normalize and impute missing metabolites
---

```{r, echo=T, message=F}


metabolite_test=as.data.frame(t(all_metabolites_raw))

#scale date, mean =0 , sd =1
scaled.test <- scale(metabolite_test)

#apply log10 transformation + scale

metabolite_test_log=log10(metabolite_test)
scaled_log_metabolites <- scale(metabolite_test_log)

#Impute na's

#Method 1: Median
impute_missing_values=function(x)

```


Adjust all ids
---

__Translate all id's to the same id (easier to deal with multiple tables)__

```{r, warning=F,message=FALSE, echo=TRUE}

#Change Metabolon ids to UMCG ids to connect later to phenotypes 
all_raw=as.data.frame(t(all_metabolites_raw))
all=as.data.frame(t(all_metabolites))

IDs_IBD=IDs_IBD[,c(1,5)]
IDs_LLD$V2=NULL

# [IMPORTANT] Table with matching ids between UMCG and Metabolon
IDs=data.frame(rbind(as.matrix(IDs_IBD), as.matrix(IDs_LLD)))
rownames(IDs)=IDs$V1
IDs$V1=NULL

# Merge to replace the ids Metabolon => UMCG
all_new_ID=merge(IDs,all, by="row.names")
all_new_ID_raw=merge(IDs,all_raw, by="row.names")

rownames(all_new_ID_raw)=all_new_ID_raw$V5
all_new_ID_raw$Row.names=NULL
all_new_ID_raw$V5=NULL
row.names(all_new_ID)=all_new_ID$V5
all_new_ID$V5=NULL
all_new_ID$Row.names=NULL
```


Data preparation and descriptive statsitics
===

Remove participants with a pouch or stoma
---

__Rational: Patients with pouch or stoma do not capture the colonic metabolite/microbiota profile__

```{r}
remove_pouch_stoma=rownames(phenos_ibd)[phenos_ibd$ibd_CurrentStomaOrPouch=="yes"]

#Keep data with stoma
all_new_ID_with_stoma=all_new_ID
all_new_ID_raw_with_stoma=all_new_ID_raw

#Remove
all_new_ID=all_new_ID[row.names(all_new_ID)%ni%remove_pouch_stoma,]
all_new_ID_raw=all_new_ID_raw[row.names(all_new_ID_raw)%ni%remove_pouch_stoma,]

```

Normalize metabolite data and split per phenotype (Control / CD / UC)
---

__Select: use inverse rank transformation or Log10__

```{r, warning=F,message=FALSE, echo=TRUE}

#Subset disease cohort
CD=rownames(phenos_ibd)[phenos_ibd$ibd_Diagnosis=="CD"]
UC=rownames(phenos_ibd)[phenos_ibd$ibd_Diagnosis=="UC"]
CT=rownames(phenos_lld)

all_new_ID_raw_cd=all_new_ID_raw[rownames(all_new_ID_raw)%in%CD,]
all_new_ID_raw_uc=all_new_ID_raw[rownames(all_new_ID_raw)%in%UC,]
all_new_ID_raw_ct=all_new_ID_raw[rownames(all_new_ID_raw)%in%CT,]

all_new_ID_cd=all_new_ID[rownames(all_new_ID)%in%CD,]
all_new_ID_uc=all_new_ID[rownames(all_new_ID)%in%UC,]
all_new_ID_ct=all_new_ID[rownames(all_new_ID)%in%CT,]

##Choose method for metabolite transformation

#Tranform to inv-rank values
#norm_metab=as.data.frame(apply(all_new_ID,2,invrank))
#norm_metab_CD=as.data.frame(apply(all_new_ID_cd,2,invrank))
#norm_metab_UC=as.data.frame(apply(all_new_ID_uc,2,invrank))
#norm_metab_CT=as.data.frame(apply(all_new_ID_ct,2,invrank))

#Use log10 transformation
norm_metab=log10(all_new_ID)
norm_metab_CD=log10(all_new_ID_cd)
norm_metab_UC=log10(all_new_ID_uc)
norm_metab_CT=log10(all_new_ID_ct)

```


Summary statistics for each metabolite
---

Important. Summary statistics are not calculated if the variance of a metabolite in the transformed table (median=1 & imputed missing values) is equal to zero, meaning that all the values are equal. This is because either all values were missing or only few samples (1 or 2) had a non-zero values that was used for imputation of the rest of the samples. 

- Calculate the number of zeros and non-zeros values on metabolites raw values

- Calculate the min, max, mean and median values on metabolites raw values

- Calculate potential outliers values (>3*IQR) on metabolites raw values 

- Calculate min and max values on transformed values (median adjusted)

- Normality test (Shapiro-test) on inverse-rank/log10 transformed data (see above). 

```{r, warning=F,message=FALSE, echo=TRUE}

summary_met=matrix(nrow=ncol(all_new_ID_raw), ncol=49)
for (i in 1:ncol(all_new_ID_raw)){
  summary_met[i,2]=sum(is.na(all_new_ID_raw[,i]))
  summary_met[i,3]=sum(!is.na(all_new_ID_raw[,i]))
  if (var(all_new_ID[,i]) == 0){
    summary_met[i,4]="NA"
    summary_met[i,5]="NA"
    summary_met[i,6]="NA"
    summary_met[i,7]="NA"
    summary_met[i,8]="NA"
    summary_met[i,9]="NA"
    summary_met[i,10]="NA"
    summary_met[i,11]="NA"
    summary_met[i,12]="NA"
    summary_met[i,13]="NA"
  }
  else{
    summary_met[i,4]=min(all_new_ID_raw[,i], na.rm = T)
    summary_met[i,5]=max(all_new_ID_raw[,i], na.rm = T)
    summary_met[i,6]=mean(all_new_ID_raw[,i], na.rm = T)
    summary_met[i,7]=median(all_new_ID_raw[,i], na.rm = T)
    summary_met[i,8]=sd(all_new_ID_raw[,i], na.rm = T)
    summary_met[i,9]=length(boxplot(all_new_ID_raw[,i], plot=FALSE, range = 3)$out)
    summary_met[i,10]=min(all_new_ID[,i], na.rm = T)
    summary_met[i,11]=max(all_new_ID[,i], na.rm = T)
    summary_met[i,12]=sd(all_new_ID[,i], na.rm = T)
    nor=shapiro.test(norm_metab[,i])
    summary_met[i,13]=nor$p.value>=0.05
  }
#Controls
  summary_met[i,14]=sum(is.na(all_new_ID_raw_ct[,i]))
  summary_met[i,15]=sum(!is.na(all_new_ID_raw_ct[,i]))
  if (var(all_new_ID_ct[,i]) == 0){
    summary_met[i,16]="NA"
    summary_met[i,17]="NA"
    summary_met[i,18]="NA"
    summary_met[i,19]="NA"
    summary_met[i,20]="NA"
    summary_met[i,21]="NA"
    summary_met[i,22]="NA"
    summary_met[i,23]="NA"
    summary_met[i,24]="NA"
    summary_met[i,25]="NA"
  }
  else{
    summary_met[i,16]=min(all_new_ID_raw_ct[,i], na.rm = T)
    summary_met[i,17]=max(all_new_ID_raw_ct[,i], na.rm = T)
    summary_met[i,18]=mean(all_new_ID_raw_ct[,i], na.rm = T)
    summary_met[i,19]=median(all_new_ID_raw_ct[,i], na.rm = T)
    summary_met[i,20]=sd(all_new_ID_raw_ct[,i], na.rm = T)
    summary_met[i,21]=length(boxplot(all_new_ID_raw_ct[,i], plot=FALSE, range = 3)$out)
    summary_met[i,22]=min(all_new_ID_ct[,i], na.rm = T)
    summary_met[i,23]=max(all_new_ID_ct[,i], na.rm = T)
    summary_met[i,24]=sd(all_new_ID_ct[,i], na.rm = T)
    nor=shapiro.test(norm_metab_CT[,i])
    summary_met[i,25]=nor$p.value>=0.05
  }
#CD
  summary_met[i,26]=sum(is.na(all_new_ID_raw_cd[,i]))
  summary_met[i,27]=sum(!is.na(all_new_ID_raw_cd[,i]))
  if (var(all_new_ID_cd[,i]) == 0){
    summary_met[i,28]="NA"
    summary_met[i,29]="NA"
    summary_met[i,30]="NA"
    summary_met[i,31]="NA"
    summary_met[i,32]="NA"
    summary_met[i,33]="NA"
    summary_met[i,34]="NA"
    summary_met[i,35]="NA"
    summary_met[i,36]="NA"
    summary_met[i,37]="NA"
  }
  else{
    summary_met[i,28]=min(all_new_ID_raw_cd[,i], na.rm = T)
    summary_met[i,29]=max(all_new_ID_raw_cd[,i], na.rm = T)
    summary_met[i,30]=mean(all_new_ID_raw_cd[,i], na.rm = T)
    summary_met[i,31]=median(all_new_ID_raw_cd[,i], na.rm = T)
    summary_met[i,32]=sd(all_new_ID_raw_cd[,i], na.rm = T)
    summary_met[i,33]=length(boxplot(all_new_ID_raw_cd[,i], plot=FALSE, range = 3)$out)
    summary_met[i,34]=min(all_new_ID_cd[,i], na.rm = T)
    summary_met[i,35]=max(all_new_ID_cd[,i], na.rm = T)
    summary_met[i,36]=sd(all_new_ID_cd[,i], na.rm = T)
    nor=shapiro.test(norm_metab_CD[,i])
    summary_met[i,37]=nor$p.value>=0.05
  }
#UC
  summary_met[i,38]=sum(is.na(all_new_ID_raw_uc[,i]))
  summary_met[i,39]=sum(!is.na(all_new_ID_raw_uc[,i]))
  if (var(all_new_ID_uc[,i]) == 0){
    summary_met[i,40]="NA"
    summary_met[i,41]="NA"
    summary_met[i,42]="NA"
    summary_met[i,43]="NA"
    summary_met[i,44]="NA"
    summary_met[i,45]="NA"
    summary_met[i,46]="NA"
    summary_met[i,47]="NA"
    summary_met[i,48]="NA"
    summary_met[i,49]="NA"
  }
  else{
    summary_met[i,40]=min(all_new_ID_raw_uc[,i], na.rm = T)
    summary_met[i,41]=max(all_new_ID_raw_uc[,i], na.rm = T)
    summary_met[i,42]=mean(all_new_ID_raw_uc[,i], na.rm = T)
    summary_met[i,43]=median(all_new_ID_raw_uc[,i], na.rm = T)
    summary_met[i,44]=sd(all_new_ID_raw_uc[,i], na.rm = T)
    summary_met[i,45]=length(boxplot(all_new_ID_raw_uc[,i], plot=FALSE, range = 3)$out)
    summary_met[i,46]=min(all_new_ID_uc[,i], na.rm = T)
    summary_met[i,47]=max(all_new_ID_uc[,i], na.rm = T)
    summary_met[i,48]=sd(all_new_ID_uc[,i], na.rm = T)
    nor=shapiro.test(norm_metab_UC[,i])
    summary_met[i,49]=nor$p.value>=0.05
  }
}
summary_met[,1]=colnames(all_new_ID_raw)
summary_met=as.data.frame(summary_met)
colnames(summary_met)=c("Metabolite", "NAs","Non_NAs", "Min", "Max", "Mean", "Median","SD", "Outliers_x3IQR", "Min_trans_data", "Max_trans_data", "SD_trans_data","Normal_distrib","NAs_CT","Non_NAs_CT", "Min_CT", "Max_CT", "Mean_CT", "Median_CT","SD_CT", "Outliers_x3IQR_CT", "Min_trans_data_CT", "Max_trans_data_CT","SD_CT", "Normal_distrib_CT", "NAs_CD","Non_NAs_CD", "Min_CD", "Max_CD", "Mean_CD", "Median_CD", "SD_CD","Outliers_x3IQR_CD", "Min_trans_data_CD", "Max_trans_data_CD", "SD_CD", "Normal_distrib_CD","NAs_UC","Non_NAs_UC", "Min_UC", "Max_UC", "Mean_UC", "Median_UC", "SD_UC","Outliers_x3IQR_UC", "Min_trans_data_UC", "Max_trans_data_UC", "SD_UC", "Normal_distrib_UC")

#Calculate presence
summary_short=summary_met[,c("Metabolite", "Non_NAs")]
summary_short2=summary_met[,c("Metabolite","Non_NAs","Non_NAs_CT","Non_NAs_CD", "Non_NAs_UC")]
summary_short2$prevalence=((as.numeric(as.character(summary_short2$Non_NAs))/nrow(all_new_ID))*100)
summary_short2$prevalence_CT=((as.numeric(as.character(summary_short2$Non_NAs_CT))/nrow(all_new_ID_ct))*100)
summary_short2$prevalence_CD=((as.numeric(as.character(summary_short2$Non_NAs_CD))/nrow(all_new_ID_cd))*100)
summary_short2$prevalence_UC=((as.numeric(as.character(summary_short2$Non_NAs_UC))/nrow(all_new_ID_raw_uc))*100)
summary_short2$Non_NAs=NULL
summary_short2$Non_NAs_CT=NULL
summary_short2$Non_NAs_CD=NULL
summary_short2$Non_NAs_UC=NULL
summary_short2[,-1] <-round(summary_short2[,-1],0)
summary_short3=melt(summary_short2)
#summary_short2=melt(summary_short, id.vars = "Metabolite")
```

Distribution of the prevalence

```{r, warning=F,message=FALSE, echo=F}
aa=ggplot(summary_short, aes(reorder(Metabolite,-as.numeric(as.character(Non_NAs))),as.numeric(as.character(Non_NAs))))+ geom_bar(color="darkblue",stat = "identity") + theme_classic() + ylab("Number of samples (n=750)") + xlab("Metabolites") +  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

ggplotly(aa)
```


Distribution of the prevalence per cohort (CT=controls, CD=Crohns, UC=Ulcerative colitis)

```{r, warning=F,message=FALSE, echo=F}
ggplot(summary_short3, aes(x=value,y=variable, fill=variable))+ geom_density_ridges_gradient(stat = "binline", bins = 20, scale=3) + theme_classic() + xlab("Present in % samples") + ylab("Number of metabolites")
```

```{r, echo=F}
summary_short4=summary_short2
rownames(summary_short4)=summary_short4$Metabolite
summary_short4$Metabolite=NULL
summary_short4=merge(annot,summary_short4, by="row.names")
rownames(summary_short4)=summary_short4$Row.names
summary_short4$Row.names=NULL
colnames(summary_short4)=c("Super pathway","Annotation", "Detected in cohort (%)", "Detected in controls (%)", "Detected in CD (%)", "Dectected in UC (%)")
datatable(summary_short4)
```


Summary statistics per sample
---


```{r, warning=F,message=FALSE, echo=TRUE}

all_new_ID_raw2=as.data.frame(t(all_new_ID_raw))
summary_sample=matrix(nrow=ncol(all_new_ID_raw2), ncol=3)

for (i in 1:ncol(all_new_ID_raw2)){
  summary_sample[i,2]=sum(is.na(all_new_ID_raw2[,i]))
  summary_sample[i,3]=sum(!is.na(all_new_ID_raw2[,i]))
}

summary_sample[,1]=colnames(all_new_ID_raw2)
summary_sample=as.data.frame(summary_sample)
colnames(summary_sample)=c("Metabolite", "NAs","Non_NAs")
summary_sample$perc_missing=((as.numeric(as.character(summary_sample$NAs))/nrow(all_new_ID_raw2))*100)
rownames(summary_sample)=summary_sample$Metabolite
summary_sample$Metabolite=NULL
summary_sample2=merge(summary_sample,cc_pheno,by="row.names")
```

Plot representing the distribution of metabolites based on its missingness 

```{r, warning=F,message=FALSE, echo=F}

cc=ggplot(summary_sample2, aes(x=perc_missing, fill=ibd_Diagnosis))+ geom_histogram(color="black" ) + theme_classic() + xlab("Percentage of missing metabolites per sample (%)") + ylab("Number of samples")

ggplotly(cc)

```


Test new vs old SCFA measurments in the population controls (4 samples with missing values)
---

```{r}

ggscatter(phenos_lld,"scfa_Butyric.acid","scfa_old_Butyrat_umol.g",add="reg.line", conf.int = T, add.params = list(color = "red",fill = "purple")) +  stat_cor( label.x = 3, method = "spearman") + xlab ("Butyric acid new measurment") + ylab ("Butyric acid old measurment")

ggscatter(phenos_lld,"scfa_Acetic.acid","scfa_old_Acetat_umol.g",add="reg.line", conf.int = T, add.params = list(color = "red",fill = "purple")) +  stat_cor( label.x = 3, method = "spearman") + xlab ("Acetate new measurment") + ylab ("Acetate old measurment")


ggscatter(phenos_lld,"scfa_Propionic.acid","scfa_old_Propionat_umol.g",add="reg.line", conf.int = T, add.params = list(color = "red",fill = "purple")) +  stat_cor( label.x = 3, method = "spearman") + xlab ("Propionate new measurment") + ylab ("Propionate old measurment")

ggscatter(phenos_lld,"scfa_Valeric.acid","scfa_old_Valerat_umol.g",add="reg.line", conf.int = T, add.params = list(color = "red",fill = "purple")) +  stat_cor( label.x = 3, method = "spearman") + xlab ("Acetate new measurment") + ylab ("Acetate old measurment")

```


Filter metabolites based on the presence / absence 
---

```{r, warning=F,message=FALSE, echo=TRUE}

to_remove_90=subset(summary_short2,prevalence_CT<10 | prevalence_CD<10 | prevalence_UC<10)

keepCT=subset(summary_short2,prevalence_CT>10)
keepCT=factor(keepCT$Metabolite)
keepCD=subset(summary_short2,prevalence_CD>10)
keepCD=factor(keepCD$Metabolite)
keepUC=subset(summary_short2,prevalence_UC>10)
keepUC=factor(keepUC$Metabolite)
```

Number of metabolites to keep after removing metabolites present in less than 10% of the samples

```{r, warning=F,message=FALSE, echo=F}
plot(euler(list(A = keepCT, B = keepCD, C = keepUC)),quantities=T, label=c("Controls","CD", "UC"))
```


```{r, warning=F,message=FALSE, echo=TRUE}
#to_remove=subset(summary_met, summary_met$perc_missing>85)
to_remove_80=subset(summary_short2,prevalence_CT<20 | prevalence_CD<20 | prevalence_CD<20)

keepCT=subset(summary_short2,prevalence_CT>20)
keepCT=factor(keepCT$Metabolite)
keepCD=subset(summary_short2,prevalence_CD>20)
keepCD=factor(keepCD$Metabolite)
keepUC=subset(summary_short2,prevalence_UC>20)
keepUC=factor(keepUC$Metabolite)
```

Numbers of metabolites to keep after removing metabolites present in less than 20% of the samples

```{r, warning=F,message=FALSE, echo=F}
plot(euler(list(A = keepCT, B = keepCD, C = keepUC)),quantities=T, label=c("Controls","CD", "UC"))
```


Removing metabolites in the normalized metabolite tables and merge phenotypes

```{r, warning=F,message=FALSE, echo=T}

# Select metabolites to remove
to_remove=subset(summary_short2,prevalence_CT<20 & prevalence_CD<20 & prevalence_UC<20)
to_remove1=to_remove
rownames(to_remove1)=to_remove1$Metabolite
to_remove1=merge(annot,to_remove1,by="row.names")
to_remove2=factor(to_remove$Metabolite)

```

```{r}
datatable(to_remove1)
```


```{r, warning=F,message=FALSE, echo=T}
#Filter/remove selected metabolites
norm_CD_filt=subset(norm_metab_CD,select = names(norm_metab_CD) %ni% to_remove2)
norm_CT_filt=subset(norm_metab_CT,select = names(norm_metab_CT) %ni% to_remove2)
norm_UC_filt=subset(norm_metab_UC,select = names(norm_metab_UC) %ni% to_remove2)
norm_filt=subset(norm_metab,select = names(norm_metab) %ni% to_remove2)
annot_v2=subset(annot,rownames(annot) %in% names(norm_filt))

#Merge phenotyes
CD_f_t=merge(cc_pheno,norm_CD_filt,by="row.names")
CT_f_t=merge(cc_pheno,norm_CT_filt,by="row.names")
UC_f_t=merge(cc_pheno,norm_UC_filt,by="row.names")
all_f_t=merge(cc_pheno,norm_filt,by="row.names")

```



Prepare taxa profiles
===

1. Focus only at species taxa level


2. Filter out taxa present in less than 20% of all samples


3. Create subsets per cohort

```{r, warning=F,message=FALSE, echo=T}
#Clean names
LLD_sp=LLD_taxa[grep("s__", rownames(LLD_taxa)), ]
IBD_sp=IBD_taxa[grep("s__", rownames(IBD_taxa)), ]
LLD_sp=LLD_sp[!grepl("t__", rownames(LLD_sp)), ]
IBD_sp=IBD_sp[!grepl("t__", rownames(IBD_sp)), ]
row.names(LLD_sp)=gsub(".*s__","",row.names(LLD_sp))
row.names(IBD_sp)=gsub(".*s__","",row.names(IBD_sp))

#Merge
taxa=as.data.frame(t(merge(IBD_sp,LLD_sp, by="row.names")))
taxa=merge(IBD_sp,LLD_sp, by="row.names")
row.names(taxa)=taxa$Row.names
taxa$Row.names=NULL
taxa=taxa[, colSums(taxa != 0) > 0]
taxa=as.data.frame(t(taxa))

#Match samples
taxa2=subset(taxa,rownames(taxa) %in% rownames(norm_filt))

#Filter present in at least 20% of the samples
taxa2=taxa2/100
taxa2[taxa2=="NA"]=0
taxa2[taxa2=="NaN"]=0
taxa_filt=taxa2[,((colSums(taxa2 !=0) / nrow(taxa2)) *100 )>20]


#Data transformation

#Choose square root arcsine transformation
taxa_filt=asin(sqrt(taxa_filt))

#Choose square root arcsine transformation
#taxa_filt2=as.data.frame(clr(as.matrix(taxa_filt)))

taxa_and_mb=merge(taxa_filt,norm_filt,by="row.names")
rownames(taxa_and_mb)=taxa_and_mb$Row.names
taxa_and_mb$Row.names=NULL

#Split per phenotype
rownames(CD_f_t)=CD_f_t$Row.names
CD_f_t$Row.names=NULL

rownames(CT_f_t)=CT_f_t$Row.names
CT_f_t$Row.names=NULL

rownames(UC_f_t)=UC_f_t$Row.names
UC_f_t$Row.names=NULL

tmb_CD=subset(taxa_and_mb,rownames(taxa_and_mb) %in% rownames(CD_f_t))
tmb_UC=subset(taxa_and_mb,rownames(taxa_and_mb) %in% rownames(UC_f_t))
tmb_CT=subset(taxa_and_mb,rownames(taxa_and_mb) %in% rownames(CT_f_t))
```


PCA: Samples clustered on metabolites
===

```{r, warning=F,message=FALSE, echo=TRUE}

pca_log= prcomp(norm_filt,scale. = T, center = T)
cmp_log=as.data.frame(pca_log$x[,1:5])
cc_pheno2=read.delim("~/Documents/Project_Metabolome/1.Input_files/mini_pheno2.txt", row.names=1)
cmp_log2=merge(cc_pheno2,cmp_log,by="row.names")
v_exp = pca_log$sdev^2
prop_v_exp= data.frame( PC=1:length(v_exp),perc= 100*(v_exp / sum(v_exp)))

```


Percentage of variance explained per component
---

```{r}
prop_v_exp2=prop_v_exp[1:20,]
ggplot(prop_v_exp2, aes (PC,perc)) + geom_bar (stat="identity") + geom_text(aes(label=round(perc, 2)), size=3, vjust=-.5) + theme_bw() + xlab ("Principal components") + ylab ("% explained variance")
```


PC1 & PC2: Diagnosis
---

```{r}
ggplot(cmp_log2, aes(PC1,-PC2, color=ibd_Diagnosis)) + geom_point() + theme_bw() + scale_colour_manual(values =c("firebrick3","black","gold3" ,"steelblue2"))
```

PC2 & PC3: Diagnosis
---

```{r}
ggplot(cmp_log2, aes(-PC2,-PC3, color=ibd_Diagnosis)) + geom_point() + theme_bw() + scale_colour_manual(values =c("firebrick3","black","gold3" ,"steelblue2"))
```

PC1 & PC2: Bowel Movements a day
---

```{r}
ggplot(cmp_log2, aes(-PC1,-PC2, color=clinical_BowelMovementADayDef)) + geom_point() + theme_bw()
```

PC1 & PC2: Bristol stool scale binary (soft vs hard)
---

```{r}
ggplot(cmp_log2, aes(-PC1,-PC2, color=clinical_BinaryBSS)) + geom_point() + theme_bw()
```

PC1 & PC2: Number of months that the sample spend in the -80 freezer
---

```{r}
ggplot(cmp_log2, aes(-PC1,-PC2, color=metabolon_Month_in_freezer)) + geom_point() + theme_bw()
```




PCA/PCoA: Microbiota composition (taxa)
===


```{r, warning=F,message=FALSE, echo=TRUE}
#Filter present in at least 20% of the samples
taxa2=subset(taxa,rownames(taxa) %in% rownames(norm_filt))
taxa2=taxa2/100
taxa2[taxa2=="NA"]=0
taxa2[taxa2=="NaN"]=0
taxa_filt=taxa2[,((colSums(taxa2 !=0) / nrow(taxa2)) *100 )>20]
#Data transformation
#Choose square root arcsine transformation
taxa_asin=asin(sqrt(taxa_filt))

#Calculate bray curtis dissimilarity
bc_dis=vegdist(taxa_asin,method = "bray")
tax_mds <- metaMDS(comm=bc_dis,k=5, autotransform = FALSE)
data.scores <- as.data.frame(scores(tax_mds))
data.scores$site <- rownames(tax_mds) 
bc_mds=merge(cc_pheno2,data.scores,by="row.names")

#Choose centered log-ratios
taxa_clr=as.data.frame(clr(as.matrix(taxa_filt)))
pca_clr= prcomp(taxa_clr,scale. = T, center = T)
cmp_clr=as.data.frame(pca_clr$x[,1:5])
cmp_clr2=merge(cc_pheno2,cmp_clr,by="row.names")
```


PCoA plot based on the Bray-Curtis dissimilarites calculated from 112 bacterial species 
---

For microbiome data we made the following selection

__a) Select only species level__
__b) Select species present in at least 20% of the samples__

```{r}

ggplot(bc_mds, aes(NMDS1,NMDS2, color=ibd_Diagnosis)) + geom_point() + theme_bw() + scale_colour_manual(values =c("firebrick3","black","gold3" ,"steelblue2"))

```


PCA plot based on the centered-log ratio abundances of 116 bacterial species
---

```{r}

ggplot(cmp_clr2, aes(PC1,PC2, color=ibd_Diagnosis)) + geom_point() + theme_bw() + scale_colour_manual(values =c("firebrick3","black","gold3" ,"steelblue2"))

```



#First cluster metabolites, then do the fit...
```{r eval=FALSE, include=FALSE}
#norm_filt_fit=subset(norm_filt,rownames(norm_filt) %in% rownames(taxa_clr))
#fit_mb_to_tx=envfit(tax_mds,norm_filt_fit)
#df <- data.frame((fit_mb_to_tx$vectors)$arrows, (fit_mb_to_tx$vectors)$r, (fit_mb_to_tx$vectors)$pvals)
#write.table(df,"~/Documents/Project_Metabolome/4.Summary_stats/envfit_test.txt", sep="\t", quote = F)
```


Explore the correlations and clustering between metabolites
====

```{r, warning=F,message=FALSE, echo=TRUE}

clustering_results=NbClust(all_new_ID, diss=NULL, distance = "euclidean", method = "ward.D2", index="all")

clustering_results=NbClust(norm_CT_filt, diss=NULL, distance = "euclidean", method = "ward.D2", index="silhouette")

```


Correlations using Spearman correlation. 

```{r, warning=F,message=FALSE, echo=TRUE}

corCD=cor(norm_CD_filt,method = "spearman")
corCT=cor(norm_CT_filt,method = "spearman")
corUC=cor(norm_UC_filt,method = "spearman")

distCD=as.dist(1-corCD)
distUC=as.dist(1-corUC)
distCT=as.dist(1-corCT)

clustCD= hclust(distCD, method = "ward.D2", members=NULL)
clustUC= hclust(distUC, method = "ward.D2", members=NULL)
clustCT= hclust(distCT, method = "ward.D2", members=NULL)


annot_v3=annot_v2
annot_v3$SUB.PATHWAY=NULL
```

Correlation plots, metabolites are ordered based on the hierarchical clustering


Metabolites in population controls

```{r, warning=F,message=FALSE, echo=F, fig.width=13, fig.height=10}
heatmaply_cor(cor(norm_CT_filt, method = "spearman"), showticklabels = c(FALSE, FALSE), main ="Metabolites in population controls", plot_method = 'plotly', scale="none", row_side_colors =annot_v3)
```


Metabolites in patients with CD


```{r, warning=F,message=FALSE, echo=F, fig.width=13, fig.height=10}
heatmaply_cor(cor(norm_CD_filt, method = "spearman"), showticklabels = c(FALSE, FALSE), main ="Metabolites in patients with CD", plot_method = 'plotly', scale="none", row_side_colors =annot_v3)
```


Metabolites in patients with UC


```{r, warning=F,message=FALSE, echo=F, fig.width=13, fig.height=10}
heatmaply_cor(cor(norm_UC_filt, method = "spearman"), showticklabels = c(FALSE, FALSE), main ="Metabolites in patients with UC", plot_method = 'plotly', scale="none", row_side_colors =annot_v3, margins=NA)
```


Heatmap including metabolites and few clinical phenotypes
---

[Subjective interpretation]

Looking at the top (horitzontal) dendogram we can see 2~4 clusters (left to right):

  1) A group of mainly patients with IBD with an enrichment of patients with CD and without ileocecal valve (due to resection). We see an enrichent of metabolites that cluster in the middle of the heatmap.
  
  2) Small cluster of patients with (mainly) UC and enrichment for stoma characterized by a lower abundance of metabolites clustering on the top part of the heatmap and higher abundance of those metabolites clustering in the middle part. 
  
  3) Clusters 3 & 4 are a mixture of patients with IBD and controls with the 4th cluster with a slightly enrichment of patients with UC and a lower abundance of the metabolites distributed in the lower section of the heatmap. 
  
  
  Diagnose: UC [n=198] / CD [n=265] / Controls [n=287]
  
  Stoma: No [n=684] / Yes [n=66]
  
  Ileocecal valve in-situ: Yes [n=331] / no [n=159] (Only cases!)
  
  Active diseases: Active [n=122] / NotActive [n=366] (Only cases! Described as combination of markers: Calprotectin levels + HBI/SSCAI)
  
  Binary bristol stool index: soft (n=207) vs hard (n=420)
  
  Bowel Movement a day: min=0 (n=57) - max=15 (n=3) / median=2 / mean=2.2 / missing=9
  
```{r, warning=F,message=FALSE, echo=TRUE}

#annotation_sel=mclean_p[,c("Diag","Box", "LC.COLUMN", "SPL.AMNT.GRAM", "RUN.DAY")]
#annotation_sel3=phenos_scfa[,c("Diag","X2.Methylbutyric.acid","Acetic.acid","Butyric.acid","Hexanoic.acid","Isobutyric.acid","Isovaleric.acid","Propionic.acid","Valeric.acid")]
#annotation_sel3[,2:9]=log10(annotation_sel3[,2:9])

annot_v3.1=all_f_t[,c("Row.names","ibd_Diagnosis","ibd_CurrentStomaOrPouch")]
rownames(annot_v3.1)=annot_v3.1$Row.names
annot_v3.1$Row.names=NULL

annot_v4=phenos_ibd[,c("ibd_Diagnosis","ibd_CurrentStomaOrPouch","ibd_ActiveDisease","ibd_ResectionAny","metabolon_month_frezzer","ibd_DiseaseLocation")]
annot_v5=phenos_lld[,c("clinical_BowelMovementADayDef","clinical_BinaryBSS","ibd_FecalCalprotectinOver200yesno","metabolon_Month_in_freezer")]
#rownames(annot_v4)=annot_v4$Row.names
#annot_v4$Row.names=NULL
```

All samples
---


```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
pheatmap(t(norm_filt), show_colnames = F,show_rownames = F, scale = "none", annotation_col = annot_v3.1, annotation_row = annot_v3, fontsize = 7, annotation_colors = list ( ibd_Diagnosis=c("CD"="firebrick2", "UC"="blue2", "Control"="white", "IBDU"="black"),ibd_CurrentStomaOrPouch=c("yes"="firebrick2","no"="white","Control"="white" ), SUPER.PATHWAY=c("Amino Acid"="lightblue", "Carbohydrate"="red", "Cofactors and Vitamins" = "gold2", "Energy"="gray79", "Lipid"="darkviolet", "Nucleotide" = "orange", "Partially Characterized Molecules" ="turquoise", "Peptide" = "limegreen", "Unknown_compount"="black", "Xenobiotics"="tan1" )))
```


Patients with CD
---


```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
annot_v4.2=annot_v4[rownames(annot_v4) %in% rownames(norm_CD_filt),]
annot_v4.2=annot_v4.2[match(rownames(norm_CD_filt), rownames(annot_v4.2)), ]

pheatmap(t(norm_CD_filt), show_colnames = F,show_rownames = F, scale = "none", annotation_col = annot_v4.2, annotation_row = annot_v3, fontsize = 7, annotation_colors = list ( ibd_Diagnosis=c("CD"="firebrick2", "UC"="blue2", "IBDU"="white"),ibd_ActiveDisease=c("Active"="firebrick2","NotActive"="white"),ibd_DiseaseLocation=c("ileum"="firebrick2","colon"="blue","both"="gray79"), ibd_ResectionAny=c("yes"="red","no"="white"), ibd_CurrentStomaOrPouch=c("yes"="red","no"="white"), SUPER.PATHWAY=c("Amino Acid"="lightblue", "Carbohydrate"="red", "Cofactors and Vitamins" = "gold2", "Energy"="gray79", "Lipid"="darkviolet", "Nucleotide" = "orange", "Partially Characterized Molecules" ="turquoise", "Peptide" = "limegreen", "Unknown_compount"="black", "Xenobiotics"="tan1" )))
```


Patients with UC
----



```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
annot_v4.2=annot_v4[rownames(annot_v4) %in% rownames(norm_UC_filt),]
annot_v4.2=annot_v4.2[match(rownames(norm_UC_filt), rownames(annot_v4.2)), ]

pheatmap(t(norm_UC_filt), show_colnames = F,show_rownames = F, scale = "none", annotation_col = annot_v4.2, annotation_row = annot_v3, fontsize = 7, annotation_colors = list ( ibd_Diagnosis=c("CD"="firebrick2", "UC"="blue2", "IBDU"="white"),ibd_ActiveDisease=c("Active"="firebrick2","NotActive"="white"),ibd_DiseaseLocation=c("ileum"="firebrick2","colon"="blue","both"="gray79"), ibd_ResectionAny=c("yes"="red","no"="white"), ibd_CurrentStomaOrPouch=c("yes"="red","no"="white"), SUPER.PATHWAY=c("Amino Acid"="lightblue", "Carbohydrate"="red", "Cofactors and Vitamins" = "gold2", "Energy"="gray79", "Lipid"="darkviolet", "Nucleotide" = "orange", "Partially Characterized Molecules" ="turquoise", "Peptide" = "limegreen", "Unknown_compount"="black", "Xenobiotics"="tan1" )))
```


Population controls
---


```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
annot_v5.2=annot_v5[rownames(annot_v5) %in% rownames(norm_CT_filt),]
annot_v5.2=annot_v5.2[match(rownames(norm_CT_filt), rownames(annot_v5.2)), ]

pheatmap(t(norm_CT_filt), show_colnames = F,show_rownames = F, scale = "none", annotation_col = annot_v5.2, annotation_row = annot_v3, fontsize = 7, annotation_colors = list ( ibd_FecalCalprotectinOver200yesno=c("yes"="firebrick2","no"="white"), SUPER.PATHWAY=c("Amino Acid"="lightblue", "Carbohydrate"="red", "Cofactors and Vitamins" = "gold2", "Energy"="gray79", "Lipid"="darkviolet", "Nucleotide" = "orange", "Partially Characterized Molecules" ="turquoise", "Peptide" = "limegreen", "Unknown_compount"="black", "Xenobiotics"="tan1" )))
```


Explorative correlation (co-abundance / co-occurance) of taxa and metabolites
---

Prepare taxonomic data: keep only species level, remove taxa present <20% of the samples, transform data using arsin square-root transformation. In total we keep 749 samples and 139 bacterial species in this step.

One IBD sample (G105251=>IBDFEC0435) does not have MGS data
 
Spearman correlation between bacteria abundances and metabolite values. Later we can consider more elegant methods.

Correlations are performed per cohort. Combining disease with control cohorts could influence the correlations (disease co-occurring events)


```{r, warning=F,message=FALSE, echo=T}

#Simple loop to correlate bateria with metabolites

pvs_cd=matrix(nrow=116, ncol=ncol(tmb_CD)-116)
rhos_cd=matrix(nrow=116, ncol=ncol(tmb_CD)-116)
x=0
for (i in 1:116){
  x=x+1
  y=0
  for (b in 117:ncol(tmb_CD)){
    y=y+1
    a=cor.test(tmb_CD[,i],tmb_CD[,b],method = "spearman")
    pvs_cd[x,y]=a$p.value
    rhos_cd[x,y]=a$estimate
  }
}
rownames(pvs_cd)=colnames(tmb_CD)[1:116]
rownames(rhos_cd)=colnames(tmb_CD)[1:116]
colnames(pvs_cd)=colnames(tmb_CD)[117:ncol(tmb_CD)]
colnames(rhos_cd)=colnames(tmb_CD)[117:ncol(tmb_CD)]

mpvs_cd=melt(pvs_cd)
mrhos_cd=melt(rhos_cd)
mpvs_cd$rho=mrhos_cd$value
mpvs_cd$fdr=p.adjust(mpvs_cd$value,method = "bonferroni")
#mlite=subset(mpvs,mpvs$fdr<0.05)


##UC


pvs_uc=matrix(nrow=116, ncol=ncol(tmb_UC)-116)
rhos_uc=matrix(nrow=116, ncol=ncol(tmb_UC)-116)
x=0
for (i in 1:116){
  x=x+1
  y=0
  for (b in 117:ncol(tmb_UC)){
    y=y+1
    a=cor.test(tmb_UC[,i],tmb_UC[,b],method = "spearman")
    pvs_uc[x,y]=a$p.value
    rhos_uc[x,y]=a$estimate
  }
}
rownames(pvs_uc)=colnames(tmb_UC)[1:116]
rownames(rhos_uc)=colnames(tmb_UC)[1:116]
colnames(pvs_uc)=colnames(tmb_UC)[117:ncol(tmb_UC)]
colnames(rhos_uc)=colnames(tmb_UC)[117:ncol(tmb_UC)]

mpvs_uc=melt(pvs_uc)
mrhos_uc=melt(rhos_uc)
mpvs_uc$rho=mrhos_uc$value
mpvs_uc$fdr=p.adjust(mpvs_uc$value,method = "bonferroni")
#mlite=subset(mpvs,mpvs$fdr<0.05)

#Controls


pvs_ct=matrix(nrow=116, ncol=ncol(tmb_CT)-116)
rhos_ct=matrix(nrow=116, ncol=ncol(tmb_CT)-116)
x=0
for (i in 1:116){
  x=x+1
  y=0
  for (b in 117:ncol(tmb_CT)){
    y=y+1
    a=cor.test(tmb_CT[,i],tmb_CT[,b],method = "spearman")
    pvs_ct[x,y]=a$p.value
    rhos_ct[x,y]=a$estimate
  }
}
rownames(pvs_ct)=colnames(tmb_CT)[1:116]
rownames(rhos_ct)=colnames(tmb_CT)[1:116]
colnames(pvs_ct)=colnames(tmb_CT)[117:ncol(tmb_CT)]
colnames(rhos_ct)=colnames(tmb_CT)[117:ncol(tmb_CT)]

mpvs_ct=melt(pvs_ct)
mrhos_ct=melt(rhos_ct)
mpvs_ct$rho=mrhos_ct$value
mpvs_ct$fdr=p.adjust(mpvs_ct$value,method = "bonferroni")
#mlite=subset(mpvs,mpvs$fdr<0.05)


```


Correlation metabolites-microbes in population controls

```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}

ab=ggplot(data = mpvs_ct, aes(Var2, Var1, fill = rho))+ geom_tile(color = "white")+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Spearman\nCorrelation\nControls") +theme_minimal()+ theme(axis.text.x=element_blank(),axis.text.y = element_text(size=5)) + ylab("Taxa") +xlab("Metabolite")

ggplotly(ab)
```


Correlation metabolites-microbes in patients with Crohn's diseases

```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
ac=ggplot(data = mpvs_cd, aes(Var2, Var1, fill = rho))+ geom_tile(color = "white")+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Spearman\nCorrelation\nCD cohort") +theme_minimal()+ theme(axis.text.x=element_blank(),axis.text.y = element_text(size=5)) + ylab("Taxa") +xlab("Metabolite")

ggplotly(ac)
```


Correlation metabolites-microbes in patients with Ulcerative colitis. 


```{r, warning=F,message=FALSE, echo=F, fig.height=15, fig.align='center',fig.width=10}
ad=ggplot(data = mpvs_uc, aes(Var2, Var1, fill = rho))+ geom_tile(color = "white")+ scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-0.7,0.7), space = "Lab", name="Spearman\nCorrelation\nUC cohort") +theme_minimal()+ theme(axis.text.x=element_blank(),axis.text.y = element_text(size=5)) + ylab("Taxa") +xlab("Metabolite")
ggplotly(ad)
```



Preliminar differential abundance analyses: diseases and host phenotypes
---
  
First we conduct analysis to identify disease trend.

__Important! Notice that here we don't take into account any counfunding or correction factors__

However, we remove patients with pouch or stomas (n=66) due to their dysbiotic profile seen at taxa and metabolic levels.

Here we simply look at the differential abundance of metabolic features comparing them between groups. We perform a Wilcoxon-test (non-parametric t-test).

Fold changes are estimated from the raw values provided by Metabolon (AUC) [log2(mean cases / mean controls)].
  
```{r, warning=F,message=FALSE, echo=T}
#mclean_raw=mclean_pre[,!colnames(mclean_pre) %in% to_remove]
m_raw_filt=subset(all_new_ID_raw,select = names(all_new_ID_raw) %ni% to_remove2)
m_raw_filt_p=merge(cc_pheno,m_raw_filt,by="row.names")

CD_controls=subset(all_f_t,all_f_t$ibd_Diagnosis!="UC")

#Remove pouches or stomas
CD_controls=subset(CD_controls,CD_controls$ibd_CurrentStomaOrPouch!="yes")
CD_controls_raw=subset(m_raw_filt_p,m_raw_filt_p$ibd_Diagnosis!="UC")

UC_controls=subset(all_f_t,all_f_t$ibd_Diagnosis!="CD")
UC_controls=subset(UC_controls,UC_controls$ibd_CurrentStomaOrPouch!="yes")
UC_controls_raw=subset(m_raw_filt_p,all_f_t$ibd_Diagnosis!="CD")


# ncol-45 (45 = number of phenotypes in the input dataframe)
cd_cnt=matrix(nrow = ncol(CD_controls)-45, ncol = 13)
x=0

for (i in 46:ncol(CD_controls)){
  x=x+1
  #cd_cnt[x,2]=median(CD_controls[,i][CD_controls$Diag=="CD"])
  cd_cnt[x,2]=median(CD_controls_raw[,i][CD_controls_raw$ibd_Diagnosis=="CD"], na.rm = T) #2
  cd_cnt[x,3]=median(UC_controls_raw[,i][UC_controls_raw$ibd_Diagnosis=="UC"], na.rm = T) #3
  #cd_cnt[x,3]=median(CD_controls[,i][CD_controls$Diag!="CD"])
  cd_cnt[x,4]=median(CD_controls_raw[,i][CD_controls_raw$ibd_Diagnosis!="CD"], na.rm = T) #4
  #cd_cnt[x,4]=mean(CD_controls[,i][CD_controls$Diag=="CD"])
  cd_cnt[x,5]=mean(CD_controls_raw[,i][CD_controls_raw$ibd_Diagnosis=="CD"], na.rm = T) #5
  cd_cnt[x,6]=mean(UC_controls_raw[,i][UC_controls_raw$ibd_Diagnosis=="UC"], na.rm = T) #6
  #cd_cnt[x,5]=mean(CD_controls[,i][CD_controls$Diag!="CD"])
  cd_cnt[x,7]=mean(CD_controls_raw[,i][CD_controls_raw$ibd_Diagnosis!="CD"], na.rm = T) #7
  #Wilcoxon effect size
  
  #Fold change
  #CD
  cd_cnt[x,8]=as.numeric(cd_cnt[x,5])/as.numeric(cd_cnt[x,7])
  cd_cnt[x,9]=log2(as.numeric(as.character(cd_cnt[x,8])))
  #UC
  cd_cnt[x,10]=as.numeric(cd_cnt[x,6])/as.numeric(cd_cnt[x,7])
  cd_cnt[x,11]=log2(as.numeric(as.character(cd_cnt[x,10])))
  
  #Wilcoxon test
  #CD
  asoc=wilcox.test(CD_controls[,i][CD_controls$ibd_Diagnosis=="CD"],CD_controls[,i][CD_controls$ibd_Diagnosis!="CD"])
  cd_cnt[x,1]=colnames(CD_controls)[i]
  cd_cnt[x,12]=asoc$p.value
  #UC
  asoc=wilcox.test(UC_controls[,i][UC_controls$ibd_Diagnosis=="UC"],UC_controls[,i][UC_controls$ibd_Diagnosis!="UC"])
  #cd_cnt[x,1]=colnames(CD_controls)[i]
  cd_cnt[x,13]=asoc$p.value
}

cd_cnt=as.data.frame(cd_cnt)

colnames(cd_cnt)=c("Metabolite", "Median_CD", "Median_UC","Median_Controls", "Mean_CD","Mean_UC", "Mean_Controls", "Fold_change_CD",  "Log2_Fold_change_CD","Fold_change_UC","Log2_Fold_change_UC", "p_value_CD", "p_value_UC")

cd_cnt$FDR_CD=p.adjust(as.numeric(as.character(cd_cnt$p_value_CD)), method = "bonferroni")
cd_cnt$FDR_UC=p.adjust(as.numeric(as.character(cd_cnt$p_value_UC)), method = "bonferroni")
cd_cnt$log10_pvalue_CD=-log10(cd_cnt$FDR_CD)
cd_cnt$log10_pvalue_UC=-log10(cd_cnt$FDR_UC)

cd_cnt2=cd_cnt
#ibd_cnt2=ibd_cnt[complete.cases(ibd_cnt$FDR),]

cd_cnt2$sig_cd="ns"
cd_cnt2$sig_uc="ns"
cd_cnt2[cd_cnt2$FDR_CD<0.05,]$sig_cd="sug"
cd_cnt2[cd_cnt2$FDR_CD<0.05 & as.numeric(as.character(cd_cnt2$Log2_Fold_change_CD))>1,]$sig_cd="up"
cd_cnt2[cd_cnt2$FDR_CD<0.05 & as.numeric(as.character(cd_cnt2$Log2_Fold_change_CD))< -1,]$sig_cd="down"
cd_cnt2[cd_cnt2$FDR_UC<0.05,]$sig_uc="sug"
cd_cnt2[cd_cnt2$FDR_UC<0.05 & as.numeric(as.character(cd_cnt2$Log2_Fold_change_UC))>1,]$sig_uc="up"
cd_cnt2[cd_cnt2$FDR_UC<0.05 & as.numeric(as.character(cd_cnt2$Log2_Fold_change_UC))< -1,]$sig_uc="down"
```

```{r, warning=F,message=FALSE, echo=T}
table(cd_cnt2$sig_cd)
table(cd_cnt2$sig_uc)
```


[WARNING, this is just an univariate analysis]

Using the above thresholds we found that almost half of the metabolites are differentially abundant on patients with CD (787/1438) and ~10% in patients with UC (158/1438). 


__Volcano plot of differentially abundant metabolites when comparing controls to patients with CD__


```{r, warning=F,message=FALSE, echo=F, fig.align='center'}
ba=ggplot(cd_cnt2,aes(x=as.numeric(as.character(Log2_Fold_change_CD)), y=log10_pvalue_CD, colour=sig_cd,text=Metabolite)) + geom_point() +
   ggtitle("Metabolites abundance CD vs Controls") +
   xlab("log2 fold change") + 
   ylab("-log10 adjusted p-value") +
   theme_bw() +
   scale_color_manual(values = c( "blue","black","gray38","coral1")) + xlim (-6,6) + geom_hline(yintercept = 1.30, linetype = "dashed", color="purple") +
   geom_vline(xintercept = c(-1, 1), linetype = "dashed", color="purple") 
ggplotly(ba)
```



__Volcano plot of differentially abundant metabolites when comparing controls to patients with UC__


```{r, warning=F,message=FALSE, echo=F, fig.align='center'}
bb=ggplot(cd_cnt2,aes(x=as.numeric(as.character(Log2_Fold_change_UC)), y=log10_pvalue_UC, colour=sig_uc,text=Metabolite)) + geom_point() +
   ggtitle("Metabolites abundance UC vs Controls") +
   xlab("log2 fold change") + 
   ylab("-log10 adjusted p-value") +
   theme_bw() +
   scale_color_manual(values = c( "blue","black","gray38","coral1")) + xlim (-6,6) + geom_hline(yintercept = 1.30, linetype = "dashed", color="purple") +
   geom_vline(xintercept = c(-1, 1), linetype = "dashed", color="purple")
ggplotly(bb)
```


Exploring potential confouders
----

__Perform a linear model including: age, sex, BMI, Bristol Stool Scale (Hard or Soft), Months that the sample spend in the freezer, bowel movements a day and diagnosis (IBD or Control)__


lm (Metabolite ~ Age + Sex + BMI + BSS + Freezer_time + Bowel movements + Diagnosis)


Testing 1368 metabolites, 


```{r, warning=F,message=FALSE, echo=T}

cc2 <- read.delim("~/Documents/Project_Metabolome/1.Input_files/mini_pheno3.txt", row.names=1)
cc2$BSS=cc2$clinical_BinaryBSS
cc2$BSS[is.na(cc2$BSS)] <- "Hard"
cc2$Diag=as.character(cc2$ibd_Diagnosis)
cc2$Diag[cc2$ibd_Diagnosis!="Control"]="IBD"
mcc2=merge(cc2,norm_filt,by="row.names")

#Test IBD vs Controls
mcc3=mcc2
#mcc3=subset(mcc2,mcc2$ibd_Diagnosis!="UC")
#mcc3=subset(mcc3,mcc3$ibd_Diagnosis!="IBDU")
tmp_results=matrix(nrow=ncol(mcc3)*7,ncol = 5)
my_col=1
for (i in 16:ncol(mcc3)){
  my_lm=lm(mcc3[,i]~host_Age_sampling+host_Sex+host_BMI+BSS+metabolon_Month_in_freezer+clinical_BowelMovementADayDef+Diag, data=mcc3)
  my_sum=summary(my_lm)
  for (a in 2:nrow(my_sum$coefficients)){
  #Metabolite
  tmp_results[my_col,1]=colnames(mcc3)[i]
  #Factor
  tmp_results[my_col,2]=rownames(my_sum$coefficients)[a]
  #Estimate
  tmp_results[my_col,3]=my_sum$coefficients[a,1]
  #SE
  tmp_results[my_col,4]=my_sum$coefficients[a,2]
  #p-value
  tmp_results[my_col,5]=my_sum$coefficients[a,4]
  my_col=my_col+1
  }
} 
my_results=as.data.frame(tmp_results)
my_results=my_results[complete.cases(my_results$V1),]
colnames(my_results)=c("Metabolite","Factor","Estimate","SE","pvalue")
my_results$FDR=p.adjust(as.numeric(as.character(my_results$pvalue)),method = "fdr")
my_sig=subset(my_results,my_results$FDR<0.05)



```


```{r, warning=F,message=FALSE, echo=T}
sig_df=as.data.frame(table(my_sig$Factor))
my_sig_met=as.data.frame(table(my_sig$Metabolite))
my_sig_met=my_sig_met[my_sig_met$Freq>0,]

BSS=subset(my_sig, my_sig$Factor=="BSSSoft")
BSS2=factor(BSS$Metabolite)

move=subset(my_sig, my_sig$Factor=="clinical_BowelMovementADayDef")
move=factor(move$Metabolite)

mon=subset(my_sig, my_sig$Factor=="metabolon_Month_in_freezer")
mon=factor(mon$Metabolite)

sex=subset(my_sig, my_sig$Factor=="host_Sexmale")
sex=factor(sex$Metabolite)


age=subset(my_sig, my_sig$Factor=="host_Age_sampling")
age=factor(age$Metabolite)


bmi=subset(my_sig, my_sig$Factor=="host_BMI")
bmi=factor(bmi$Metabolite)


ibd=subset(my_sig, my_sig$Factor=="DiagIBD")
ibd=factor(ibd$Metabolite)



my_test_list=list(IBD =ibd, BMI = bmi, Age = age, Sex=sex ,Month_in_freezer=mon, Bowel_move=move, BSS=BSS2)

```

__Number of associated metabolites per feature__


```{r}
datatable(sig_df)
```


IBD and Bristol stool scale are the parameters that contribute to the largest (and significant) proportion of metabolites variation. However, a proportion of the metabolites (~100/1368 is associated to the time that the sample was stored in a freezer -80C), showing the importance of potential technical confounders in the further associations analyses. 


```{r, warning=F,message=FALSE, echo=F}
upset(fromList(my_test_list), order.by = "freq", nsets = 7)
```


